---
title: "Hands-on Exercise 6: Geographical Segmentation with Spatially Constrained Clustering Techniques"
author: "Pwint Phoo Thaw"
date-modified: "last-modified"
execute:
  echo: true
  eva: true
  warning: false
  freeze: true
---

## **1.1 Overview**

This hands-on experience is on how to delineate homogeneous region by using geographically referenced multivariate data. There are two major analysis, namely:

-   hierarchical cluster analysis; and

-   spatially constrained cluster analysis.

### **1.1.1 Learning Outcome**

By the end of this hands-on exercise, you will able:

-   to convert GIS polygon data into R’s simple feature data.frame by using appropriate functions of **sf** package of R;

-   to convert simple feature data.frame into R’s SpatialPolygonDataFrame object by using appropriate **sf** of package of R;

-   to perform custer analysis by using *hclust()* of Base R;

-   to perform spatially constrained cluster analysis using *skater()* of Base R; and

-   to visualise the analysis output by using **ggplot2** and **tmap** package.

## **1.2 Getting Started**

### **1.2.1 The analytical question**

In geobusiness and spatial policy, it is a common practice to delineate the market or planning area into homogeneous regions by using multivariate data. In this hands-on exercise, we are interested to delineate [Shan State](https://en.wikipedia.org/wiki/Shan_State), [Myanmar](https://en.wikipedia.org/wiki/Myanmar) into homogeneous regions by using multiple Information and Communication technology (ICT) measures, namely: Radio, Television, Land line phone, Mobile phone, Computer, and Internet at home.

## **1.3 The data**

Two data sets will be used in this study. They are:

-   Myanmar Township Boundary Data (i.e. *myanmar_township_boundaries*) : This is a GIS data in ESRI shapefile format. It consists of township boundary information of Myanmar. The spatial data are captured in polygon features.

-   *Shan-ICT.csv*: This is an extract of [**The 2014 Myanmar Population and Housing Census Myanmar**](https://myanmar.unfpa.org/en/publications/2014-population-and-housing-census-myanmar-data-sheet) at the township level.

Both data sets are download from [Myanmar Information Management Unit (MIMU)](http://themimu.info/)

### **1.3.1 Installing and loading R packages**

-   Cluster analysis

    -   **cluster**

    -   **ClustGeo**

```{r}
pacman::p_load(spdep, tmap, sf, ClustGeo, 
               ggpubr, cluster, factoextra, NbClust,
               heatmaply, corrplot, psych, tidyverse, GGally)
```

## **1.4 Data Import and Prepatation**

### **1.4.1 Importing geospatial data into R environment**

In this section, Myanmar Township Boundary GIS data will be imported and its associated attrbiute table into R environment.

```{r}
shan_sf <- st_read(dsn = "data/geospatial", 
                   layer = "myanmar_township_boundaries") %>%
  filter(ST %in% c("Shan (East)", "Shan (North)", "Shan (South)")) %>%
  select(c(2:7))
```

```{r}
shan_sf
```

```{r}
glimpse(shan_sf)
```

### **1.4.2 Importing aspatial data into R environment**

The csv file will be import using *read_csv* function of **readr** package.

```{r}
ict <- read_csv ("data/aspatial/Shan-ICT.csv")
```

```{r}
summary(ict)
```

### **1.4.3 Derive new variables using dplyr package**

The unit of measurement of the values are number of household. Using these values directly will be bias by the underlying total number of households. In general, the townships with relatively higher total number of households will also have higher number of households owning radio, TV, etc.

In order to overcome this problem, we will derive the penetration rate of each ICT variable by using the code chunk below.

```{r}
ict_derived <- ict %>%
  mutate(`RADIO_PR` = `Radio`/`Total households`*1000) %>%
  mutate(`TV_PR` = `Television`/`Total households`*1000) %>%
  mutate(`LLPHONE_PR` = `Land line phone`/`Total households`*1000) %>%
  mutate(`MPHONE_PR` = `Mobile phone`/`Total households`*1000) %>%
  mutate(`COMPUTER_PR` = `Computer`/`Total households`*1000) %>%
  mutate(`INTERNET_PR` = `Internet at home`/`Total households`*1000) %>%
  rename(`DT_PCODE` =`District Pcode`,`DT`=`District Name`,
         `TS_PCODE`=`Township Pcode`, `TS`=`Township Name`,
         `TT_HOUSEHOLDS`=`Total households`,
         `RADIO`=`Radio`, `TV`=`Television`, 
         `LLPHONE`=`Land line phone`, `MPHONE`=`Mobile phone`,
         `COMPUTER`=`Computer`, `INTERNET`=`Internet at home`) 
```

```{r}
summary(ict_derived)
```

## **1.5 Exploratory Data Analysis (EDA)**

### **1.5.1 EDA using statistical graphics**

We can plot the distribution of the variables (i.e. Number of households with radio) by using appropriate Exploratory Data Analysis (EDA) as shown in the code chunk below.

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

Boxplot is useful to detect if there are outliers.

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO`)) +
  geom_boxplot(color="black", 
               fill="light blue")

```

Next, we will also plotting the distribution of the newly derived variables (i.e. Radio penetration rate) by using the code chunk below.

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO_PR`)) +
  geom_boxplot(color="black", 
               fill="light blue")
```

### For **RADIO_PR** (processed/normalized values)

-   **Histogram** : Distribution looks more balanced and nearly symmetric. Values are spread fairly evenly from 0 to 500, with many observations clustering between 150–250. No extreme concentrations at the lower end.

-   **Boxplot** : Displays a more symmetric spread with a wide interquartile range. Only one mild outlier (\~500). The median is centered, indicating little skewness.

-   **Interpretation**: After transformation, RADIO_PR is much more normalized, balanced, and less influenced by extreme values. The skewness and heavy tails observed in RADIO are largely corrected.

```         
```

### 1.5.2 EDA using choropleth map

#### 1.5.2.1 Joining geospatial data with aspatial data

```{r}

shan_sf <- left_join(shan_sf, ict_derived, by = c("TS_PCODE" = "TS_PCODE")) 

write_rds(shan_sf, "data/rds/shan_sf.rds")

```

```{r}




shan_sf <- read_rds("data/rds/shan_sf.rds") 
```

#### 1.5.2.2 Preparing a choropleth map

```{r}


qtm(shan_sf, "RADIO_PR")      
```

```{r}
TT_HOUSEHOLDS.map <- tm_shape(shan_sf) + 
  tm_fill(col = "TT_HOUSEHOLDS",
          n = 5,
          style = "jenks", 
          title = "Total households") + 
  tm_borders(alpha = 0.5) 

RADIO.map <- tm_shape(shan_sf) + 
  tm_fill(col = "RADIO",
          n = 5,
          style = "jenks",
          title = "Number Radio ") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(TT_HOUSEHOLDS.map, RADIO.map,
             asp=NA, ncol=2)

```

```{r}

tm_shape(shan_sf) +
    tm_polygons(c("TT_HOUSEHOLDS", "RADIO_PR"),
                style="jenks") +
    tm_facets(sync = TRUE, ncol = 2) +
  tm_legend(legend.position = c("right", "bottom"))+
  tm_layout(outer.margins=0, asp=0)

```

## 1.6 Correlation Analysis

Before we perform cluster analysis, it is important for us to ensure that the cluster variables are not highly correlated.

In this section, you will learn how to use [*corrplot.mixed()*](https://cran.r-project.org/web/packages/corrplot/corrplot.pdf) function of [**corrplot**](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html) package to visualise and analyse the correlation of the input variables.

```{r}

cluster_vars.cor = cor(ict_derived[,12:17])
corrplot.mixed(cluster_vars.cor,
         lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")

```

> The plot usually shows **COMPUTER_PR** and **INTERNET_PR** as highly correlated; we will keep only one (COMPUTER_PR) for clustering to avoid redundancy.

## 1.7 Hierarchy Cluster Analysis

In this section, you will learn how to perform hierarchical cluster analysis. The analysis consists of four major steps:

### 1.7.1 Extracting clustering variables

```{r}

cluster_vars <- shan_sf %>%
  st_set_geometry(NULL) %>%
  select("TS.x", "RADIO_PR", "TV_PR", "LLPHONE_PR", "MPHONE_PR", "COMPUTER_PR")
head(cluster_vars,10)
```

Notice that the final clustering variables list does not include variable INTERNET_PR because it is highly correlated with variable COMPUTER_PR.

Next, we need to change the rows by township name instead of row number by using the code chunk below

```{r}

row.names(cluster_vars) <- cluster_vars$"TS.x"
head(cluster_vars,10)

```

Notice that the row number has been replaced into the township name.

Now, we will delete the TS.x field by using the code chunk below.

```{r}
shan_ict <- select(cluster_vars, c(2:6))
head(shan_ict, 10)

```

```{r}

shan_ict <- read_rds("data/rds/shan_ict.rds")

```

### 1.7.2 Data Standardisation

Multiple variables have different ranges; we standardise before computing distances.

### 1.7.3 Min--Max standardisation

```{r}

shan_ict.std <- normalize(shan_ict)
summary(shan_ict.std)

```

Notice that the values range of the Min-max standardised clustering variables are 0-1 now.

### 1.7.4 Z-score standardisation

```{r}
shan_ict.z <- scale(shan_ict)
describe(shan_ict.z)

```

### 12.7.5 Visualising the standardised variables

Beside reviewing the summary statistics of the standardised clustering variables, it is also a good practice to visualise their distribution graphical.

The code chunk below plot the scaled ***Radio_PR*** field.

```{r}
r <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Raw values without standardisation")

shan_ict_s_df <- as.data.frame(shan_ict.std)
s <- ggplot(data=shan_ict_s_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Min-Max Standardisation")

shan_ict_z_df <- as.data.frame(shan_ict.z)
z <- ggplot(data=shan_ict_z_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Z-score Standardisation")

ggarrange(r, s, z,
          ncol = 3,
          nrow = 1)
```

###### What statistical conclusion can you draw from the histograms above?

The histograms highlight that standardisation mainly adjusts the scale of the data while leaving the overall distribution shape largely unchanged. Keeping raw values can work when all features are measured on the same scale, but if variables differ in magnitude, larger ones may dominate and skew the results, especially in distance-based or optimization methods. Min–Max scaling brings values into a fixed \[0,1\] range and preserves proportionality, though it can be heavily influenced by outliers. Z-score standardisation, which centers the data around zero with unit variance, is generally more reliable for comparing across different units and is often preferred for methods that assume standardized inputs, such as regression, PCA, or clustering. In practice, the method you choose should depend on the type of data, the presence of outliers, and the requirements of the analysis technique.

```{r}

r <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_density(color="black",
               fill="light blue") +
  ggtitle("Raw values without standardisation")

shan_ict_s_df <- as.data.frame(shan_ict.std)
s <- ggplot(data=shan_ict_s_df, 
       aes(x=`RADIO_PR`)) +
  geom_density(color="black",
               fill="light blue") +
  ggtitle("Min-Max Standardisation")

shan_ict_z_df <- as.data.frame(shan_ict.z)
z <- ggplot(data=shan_ict_z_df, 
       aes(x=`RADIO_PR`)) +
  geom_density(color="black",
               fill="light blue") +
  ggtitle("Z-score Standardisation")

ggarrange(r, s, z,
          ncol = 3,
          nrow = 1)

```

### 1.7.6 Computing proximity matrix

In R, many packages provide functions to calculate distance matrix. We will compute the proximity matrix by using dist() of R.

`dist()` supports six distance proximity calculations, they are: **euclidean**, **maximum**, **manhattan**, **canberra**, **binary and minkowski**. The default is ***euclidean*** proximity matrix.

The code chunk below is used to compute the proximity matrix using euclidean method.

```{r}


proxmat <- stats::dist(shan_ict, method = 'euclidean') 

proxmat

```

### 1.7.7 Computing hierarchical clustering

```{r}


hclust_ward <- hclust(proxmat, method = 'ward.D')  

plot(hclust_ward, cex = 0.6) 

```

### 1.7.8 Selecting the optimal clustering algorithm

One of the challenge in performing hierarchical clustering is to identify stronger clustering structures. The issue can be solved by using use `agnes()` function of cluster package. It functions like `hclus()`, however, with the `agnes()` function we can also get the agglomerative coefficient, which measures the amount of clustering structure found (values closer to 1 suggest strong clustering structure).

The code chunk below will be used to compute the agglomerative coefficients of all hierarchical clustering algorithms.

```{r}
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

ac <- function(x) {
  agnes(shan_ict, method = x)$ac
}

map_dbl(m, ac)

```

### 1.7.9 Determining Optimal Clusters

Another technical challenge face by data analyst in performing clustering analysis is to determine the optimal clusters to retain.

There are **three** commonly used methods to determine the optimal clusters, they are:

-   [Elbow Method](https://en.wikipedia.org/wiki/Elbow_method_(clustering))\
-   [Average Silhouette Method](https://www.sciencedirect.com/science/article/pii/0377042787901257?via%3Dihub)\
-   [Gap Statistic Method](http://www.web.stanford.edu/~hastie/Papers/gap.pdf)

#### 1.7.9.1 Gap Statistic Method

The [gap statistic](http://www.web.stanford.edu/~hastie/Papers/gap.pdf) compares the total within intra-cluster variation for different values of k with their expected values under null reference distribution of the data. The estimate of the optimal clusters will be value that maximize the gap statistic (i.e., that yields the largest gap statistic). This means that the clustering structure is far away from the random uniform distribution of points.

To compute the gap statistic, `clusGap()` of **cluster package** will be used.

```{r}
set.seed(12345)
gap_stat <- clusGap(shan_ict, 
                    FUN = hcut, 
                    nstart = 25, 
                    K.max = 10, 
                    B = 50)
# Print the result
print(gap_stat, method = "firstmax")

```

Also note that the `hcut` function used is from factoextra package.

Next, we can visualise the plot by using `fviz_gap_stat()` of **factoextra package**.fviz_gap_stat(gap_stat)

```{r}
fviz_gap_stat(gap_stat)


```

With reference to the gap statistic graph above, the recommended number of cluster to retain is 1. However, it is not logical to retain only one cluster. By examine the gap statistic graph, the 6-cluster gives the largest gap statistic and should be the next best cluster to pick.

### 12.7.10 Interpreting the dendrograms

In the dendrogram displayed above, each leaf corresponds to one observation. As we move up the tree, observations that are similar to each other are combined into branches, which are themselves fused at a higher height.

The height of the fusion, provided on the vertical axis, indicates the (dis)similarity between two observations. The higher the height of the fusion, the less similar the observations are. Note that, conclusions about the proximity of two observations can be drawn only based on the height where branches containing those two observations first are fused. We cannot use the proximity of two observations along the horizontal axis as a criteria of their similarity.

It’s also possible to draw the dendrogram with a border around the selected clusters by using `rect.hclust()` of R stats. The argument border is used to specify the border colors for the rectangles.

```{r}
plot(hclust_ward, cex = 0.6)
rect.hclust(hclust_ward, 
            k = 6, 
            border = 2:5)

```

### 1.7.11 Visually-driven hierarchical clustering analysis

```{r}


shan_ict_mat <- data.matrix(shan_ict) 
heatmaply(normalize(shan_ict_mat),
          Colv=NA,
          dist_method = "euclidean",
          hclust_method = "ward.D",
          seriate = "OLO",
          colors = Blues,
          k_row = 6,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="Geographic Segmentation of Shan State by ICT indicators",
          xlab = "ICT Indicators",
          ylab = "Townships of Shan State"
          )

```

### 1.7.12 Mapping the clusters formed

```{r}

groups <- as.factor(cutree(hclust_ward, k=6))

```

```{r}
shan_sf_cluster <- cbind(shan_sf, as.matrix(groups)) %>%
  rename(`CLUSTER`=`as.matrix.groups.`)

```

```{r}

qtm(shan_sf_cluster, "CLUSTER")

```

## 1.8 Spatially Constrained Clustering: SKATER approach

In this section, we will learn how to derive spatially constrained cluster by using `skater()` method of **spdep package**.

### 1.8.1 Converting into SpatialPolygonsDataFrame

```{r}


shan_sp <- sf::as_Spatial(shan_sf) 
```

### 1.8.2 Computing Neighbour List

```{r}
shan.nb <- poly2nb(shan_sp)
summary(shan.nb)
```

```{r}

coords <- st_coordinates(
  st_centroid(st_geometry(shan_sf)))
```

```{r}
plot(st_geometry(shan_sf), 
     border=grey(.5))
plot(shan.nb,
     coords, 
     col="blue", 
     add=TRUE)

```

### 1.8.3 Computing minimum spanning tree — edge costs

#### 1.8.3.1 Calculating edge costs

```{r}

lcosts <- nbcosts(shan.nb, shan_ict)

```

```{r}


shan.w <- spdep::nb2listw(shan.nb, lcosts, style = "B")
summary(shan.w)

```

### 1.8.4 Computing minimum spanning tree

```{r}


shan.mst <- spdep::mstree(shan.w) 

```

```{r}


class(shan.mst); 

```

```{r}

dim(shan.mst)

```

> The dimension is 54 and not 55. This is because the minimum spanning tree consists on n-1 edges (links) in order to traverse all the nodes.

```{r}

# Inspect the content of shan.mst
head(shan.mst)

```

```{r}
plot(st_geometry(shan_sf), 
                 border=gray(.5))
plot.mst(shan.mst, 
         coords, 
         col="blue", 
         cex.lab=0.7, 
         cex.circles=0.005, 
         add=TRUE)
```

### 1.8.5 Computing spatially constrained clusters using SKATER method

```{r}
clust6 <- spdep::skater(edges = shan.mst[,1:2], 
                 data = shan_ict, 
                 method = "euclidean", 
                 ncuts = 5)

```

```{r}

str(clust6) 

```

```{r}


cc6 <- clust6$groups; cc6 
cc6

```

We can find out how many observations are in each cluster by means of the table command. Parenthetially, we can also find this as the dimension of each vector in the lists contained in edges.groups. For example, the first list has node with dimension 12, which is also the number of observations in the first cluster.

```{r}

table(cc6)

```

Lastly, we can also plot the pruned tree that shows the five clusters on top of the townshop area.

```{r}
plot(st_geometry(shan_sf), 
     border=gray(.5))
plot(clust6, 
     coords, 
     cex.lab=.7,
     groups.colors=c("red","green","blue", "brown", "pink"),
     cex.circles=0.005, 
     add=TRUE)

```

### 1.8.6 Visualising the clusters in choropleth map

The code chunk below is used to plot the newly derived clusters by using SKATER method.

```{r}

groups_mat <- as.matrix(clust6$groups)
shan_sf_spatialcluster <- cbind(shan_sf_cluster, as.factor(groups_mat)) %>%
  rename(`SP_CLUSTER`=`as.factor.groups_mat.`)
qtm(shan_sf_spatialcluster, "SP_CLUSTER")
```

For easy comparison, it will be better to place both the hierarchical clustering and spatially constrained hierarchical clustering maps next to each other.

```{r}
hclust.map <- qtm(shan_sf_cluster,
                  "CLUSTER") + 
  tm_borders(alpha = 0.5) 

shclust.map <- qtm(shan_sf_spatialcluster,
                   "SP_CLUSTER") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(hclust.map, shclust.map,
             asp=NA, ncol=2)

```

## 1.9 Spatially Constrained Clustering: ClustGeo Method

### 1.9.1 A short note about ClustGeo

**ClustGeo** implements Ward‑like hierarchical clustering with a mixing parameter **alpha ∈ \[0,1\]** combining attribute dissimilarities (**D0**) and spatial dissimilarities (**D1**). Use **choicealpha()** to pick alpha balancing contiguity and attribute fit.

### 1.9.2 Ward‑like hierarchical clustering (non‑spatial): ClustGeo

```{r}
nongeo_cluster <- hclustgeo(proxmat)
plot(nongeo_cluster, cex = 0.5)
rect.hclust(nongeo_cluster, 
            k = 6, 
            border = 2:5)
```

#### 12.9.2.1 Mapping the clusters formed

Similarly, we can plot the clusters on a categorical area shaded map by using the steps we learned in 5.7.12 Mapping the clusters formed.

```{r}
groups <- as.factor(cutree(nongeo_cluster, k=6))
```

```{r}
shan_sf_ngeo_cluster <- cbind(shan_sf, as.matrix(groups)) %>%
  rename(`CLUSTER` = `as.matrix.groups.`)
```

```{r}
qtm(shan_sf_ngeo_cluster, "CLUSTER")
```

### 1.9.3 Spatially Constrained Hierarchical Clustering

```{r}

dist <- sf::st_distance(shan_sf, shan_sf) 


distmat <- stats::as.dist(dist)

```

```{r}
cr <- choicealpha(proxmat, distmat, range.alpha = seq(0, 1, 0.1), K=6, graph = TRUE)

```

```{r}
clustG <- hclustgeo(proxmat, distmat, alpha = 0.2)

```

```{r}


groups <- as.factor(cutree(clustG, k = 6)) 

shan_sf_Gcluster <- cbind(shan_sf, as.matrix(groups)) %>% 
dplyr::rename(`CLUSTER` = `as.matrix.groups.`)


qtm(shan_sf_Gcluster, "CLUSTER") 

```

## 1.10 Visual Interpretation of Clusters

### 1.10.1 Visualising individual clustering variable

```{r}

ggplot(data = shan_sf_ngeo_cluster,
       aes(x = CLUSTER, y = RADIO_PR)) +
  geom_boxplot()

```

> The boxplot reveals Cluster 3 displays the highest mean Radio Ownership Per Thousand Household. This is followed by Cluster 2, 1, 4, 6 and 5.

### 1.10.2 Multivariate Visualisation

```{r}

ggparcoord(data = shan_sf_ngeo_cluster, 
           columns = c(17:21), 
           scale = "globalminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of ICT Variables by Cluster") +
  facet_grid(~ CLUSTER) + 
  theme(axis.text.x = element_text(angle = 30))
```

The parallel coordinate plot above reveals that households in Cluster 4 townships tend to own the highest number of TV and mobile-phone. On the other hand, households in Cluster 5 tends to own the lowest of all the five ICT.

Note that the **scale** argument of `ggparcoor()` provide several methods to scale the clustering variables. They are:

-   std: univariately, subtract mean and divide by standard deviation.\
-   robust: univariately, subtract median and divide by median absolute deviation.\
-   uniminmax: univariately, scale so the minimum of the variable is zero, and the maximum is one.\
-   globalminmax: no scaling is done; the range of the graphs is defined by the global minimum and the global maximum.\
-   center: use uniminmax to standardize vertical height, then center each variable at a value specified by the scaleSummary param.\
-   centerObs: use uniminmax to standardize vertical height, then center each variable at the value of the observation specified by the centerObsID param

There is no one best scaling method to use. You should explore them and select the one that best meet your analysis need.

Last but not least, we can also compute the summary statistics such as mean, median, sd, etc to complement the visual interpretation.

In the code chunk below, `group_by()` and `summarise()` of dplyr are used to derive mean values of the clustering variables.

```{r}
shan_sf_ngeo_cluster %>% 
  st_set_geometry(NULL) %>%
  group_by(CLUSTER) %>%
  summarise(mean_RADIO_PR = mean(RADIO_PR),
            mean_TV_PR = mean(TV_PR),
            mean_LLPHONE_PR = mean(LLPHONE_PR),
            mean_MPHONE_PR = mean(MPHONE_PR),
            mean_COMPUTER_PR = mean(COMPUTER_PR))
```
