---
title: "Take-Home Exercise 1: The Influx of Chinese Bubble Tea Brands in Singapore: A Spatio-Temporal Geospatial Analysis (2024–2025)"
author: "Pwint Phoo Thaw"
date: "September 12, 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **Problem Statement**

Between **2024 and mid-2025**, Singapore has experienced an influx of **popular Chinese bubble tea brands** such as *Mixue, HEYTEA, and Chagee*. These entrants are competing directly with established Taiwanese and local players (*KOI, LiHo, Gong Cha*), reshaping the city’s retail food and beverage landscape.

Understanding where these Chinese brands are locating, when they enter, and how they interact with existing bubble tea ecosystems is crucial for urban retail planning, zoning policies, and SME competitiveness.

This study applies geospatial analytics to uncover the **spatial clustering, temporal waves of entry, and proximity to key infrastructure (MRTs, malls, schools)** of Chinese bubble tea shops, compared against incumbents. The findings will inform **planning authorities, mall operators, and policy makers** on how foreign retail entrants shape Singapore’s urban form and consumer spaces.

## **1. Objectives**

-   **Map Spatio-Temporal Distribution**\
    Track when and where Chinese bubble tea brands entered (2024–Jun 2025), and compare hotspots with existing players.

-   **Assess Clustering & Competition**\
    Use spatial statistics (KDE, Ripley’s K, Cross-K) to test clustering and co-location with incumbents.

-   **Identify Temporal Waves**\
    Apply spatio-temporal KDE (STKDE) to reveal expansion waves from CBD to heartlands.

-   **Examine Infrastructure Proximity**\
    Test whether outlets concentrate near MRTs, malls, and schools.

-   **Derive Policy Insights**\
    Translate findings into implications for retail zoning, transit-oriented planning, and SME support.

## **2. Data Wrangling**

```{r}

pacman::p_load(readr, dplyr, tidyr, stringr, lubridate,terra,spatstat,
               sf, tmap, units, httr, jsonlite, tidyverse,rvest )

```

**2.1 Importing ACRA Data (Aspatial)**

```{r}
folder_path <- "data/aspatial/acra"
file_list <- list.files(path=folder_path,
                        pattern = "^ACRA*.*\\.csv$",
                        full.names = TRUE)

acra_data <- file_list %>%
  map_dfr(read_csv)
```

**Saving ACRA data**

```{r}
write_rds(acra_data,
          "data/rds/acra_data.rds") 
```

**Tidying ACRA data (56123 is for bubble tea shops)**

```{r}
### select target businesess, derive year, tidy postal code for geo coding and...
biz_56123 <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code == 56123) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date, label = TRUE, abbr = TRUE)) %>%
  mutate(postal_code = str_pad(postal_code, width = 6, side = "left", pad = "0")) %>%
   filter((YEAR == 2024 & MONTH_NUM >= 1) | (YEAR == 2025 & MONTH_NUM <= 6))

```

```{r}

#|eval:false
#|echo:false
write_rds(biz_56123,
"data/rds/biz_56123.rds")

```

```{r}
biz_56123 <-
read_rds("data/rds/biz_56123.rds")



```

```{r}
#biz_56123 <- read_rds("data/rds/biz_56123.rds")

#par(mar = c(4, 4, 2, 1))  # set smaller margins
# plot(biz_56123)  # uncomment if it's an sf object and you want a quick plot

```

**Geocoding (for finding X,Y coordinates)**

```{r}
postcodes <- unique(biz_56123$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"  ##API without paying

found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )


###converting to data frame
res <- GET(url, query = query)
json <- content(res)

if (json$found != 0) {
  df <- as.data.frame(json$results, stringsAsFactors = FALSE)
  df$input_postcode <- pc
  found <- bind_rows(found, df)
  } else {
  not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
}

```

```{r}
found <- found %>%
  select(1:10)
```

```{r}

```

Appending the location information (join it back with 56123 right one with new geo_coded postal_code

2 columns in meter, 2 columns in Lon and Lat

```{r}
biz_56123_join = biz_56123 %>%
   left_join(found,
             by = c('postal_code' = 'POSTAL'))
```

```{r}
write_rds(biz_56123_join, "data/rds/biz_56123_join.rds")
```

**Converting into SF data frame**

```{r}
biz_56123_sf <- st_as_sf(biz_56123_join, 
                         coords = c("X","Y"),
                         crs=3414) 
class(biz_56123_sf)
plot(st_geometry(biz_56123_sf))
```

```{r}
# case-insensitive regex for trailing street-type words (and anything after them)
suffix_re <- "(?i)\\s+(ROAD|RD|STREET|ST|AVENUE|AVE|DRIVE|DR|LANE|LN|CLOSE|CL|CRESCENT|CRES|TERRACE|TERR|WALK|PLACE|PL|PARK|PK|HEIGHTS|HTS|LINK|VIEW|WAY|COURT|CT|CIRCUIT|CIR|RISE|GROVE|GARDENS|GDN)(\\s+.*)?$"

# create Zone from street_name (fallback to STREET_NAME if needed)
biz_56123_sf <- biz_56123_sf %>%
  mutate(
    street_raw = coalesce(street_name),
    Zone = street_raw %>%
      str_replace_all("_", " ") %>%  # tidy underscores
      str_squish() %>%               # trim extra spaces
      str_remove(suffix_re) %>%      # drop trailing street-type
      str_squish() %>%
      str_to_upper()                 # <-- UPPER CASE
  )

# quick check
biz_56123_sf %>% select(street_raw, Zone) %>% head()
```

```{r}
write_rds(biz_56123_sf, "data/rds/biz_56123_sf.rds")
```

**2.2 Importing Geospatial Data**

```{r}
mpsz_sf <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>%
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
mpsz_sf <- mpsz_sf %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
###testing hands on 2.2
mpsz_cl <- mpsz_sf %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS",
         PLN_AREA_N != "NORTH-EASTERN ISLANDS")
```

```{r}
write_rds(mpsz_cl, 
          "data/rds/mpsz_cl.rds")
```

```{r}
st_crs(mpsz_cl) == st_crs(biz_56123_sf)
```

```{r}
ggplot() +
  geom_sf(data = mpsz_cl, fill = "gray", color = "black", size = 0.2) +
  geom_sf(data = biz_56123_sf, color = "black", size = 1.2, alpha = 0.7) +
  labs(title = "Food kiosks Across Singapore Subzones",
       subtitle = "All layers aligned to EPSG:3414 (SVY21)",
       caption = "Source: URA Master Plan & Data.gov.sg") +
  theme_minimal()
```

**2.3 Joining ARCA data and Geospatial data**

```{r}
#####will do later
```

**Objective 1:**

```{r}
biz_56123_ppp <- as.ppp(biz_56123_sf)

```

```{r}
class(biz_56123_ppp)
```

```{r}
summary(biz_56123_ppp)
```

#### **Creating *owin* object and combining point events**

```{r}
sg_owin <- as.owin(mpsz_cl)
class(sg_owin)
```

```{r}
biz56123_sg_ppp = biz_56123_ppp [sg_owin]
```

#### **Clark-Evan Test for Nearest Neighbour Analysis**

```{r}
clarkevans.test(biz56123_sg_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```

### **1) Statistical conclusion**

-   The test statistic **R = 0.514** (well **below 1**) and the **p-value \< 2.2e-16**.

-   Translation: the food kioks locations are **much closer together than we’d expect by random chance**—i.e., they are **strongly clustered**, not evenly spread out.
