---
title: "Take-Home Exercise 1: Hotspots on the Go: Clustering of Food Kiosks Across Singapore, 2024–Jun 2025"
author: "Pwint Phoo Thaw"
date: "September 12, 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **Problem Statement**

From **Jan 2024 to Jun 2025**, Singapore saw rapid growth of **food kiosks (SSIC 56123)**—small, takeaway-focused outlets with little or no seating. Their expansion around **MRT nodes, malls, schools, offices, and HDB towns** can reshape micro-retail patterns, affect pedestrian space and curb logistics, and create pockets of under- or over-supply. This study uses ACRA registrations to reveal **where** and **when** kiosks emerge, how strongly they **cluster**, how activity **evolves over time**, and how locations relate to key **urban infrastructure**, to inform **evidence-based retail and land-use planning** in Singapore.

## **1. Objectives**

-   **Map Spatio-Temporal Distribution**\
    Track when and where new 56123 kiosks entered (2024–Jun 2025) and highlight hotspot areas by month and subzone.
-   **Assess Clustering \***\
    Use spatial statistics (KDE, Ripley’s K/L or O-ring) to test whether kiosks are clustered or dispersed and at what spatial scales.
-   **Identify Temporal Waves**\
    Apply spatio-temporal KDE (STKDE) to reveal waves of kiosk entry (e.g., CBD/malls first, then diffusion to heartlands).
-   **Examine Infrastructure Proximity \***\
    Test whether kiosks concentrate near **MRT stations, malls, and schools** (buffer analysis; simple statistical tests).
-   **Derive Planning Insights**\
    Translate findings into implications for **zoning/tenant-mix**, **transit-oriented clustering management**, curb/logistics planning, and **SME support**.

## **2. Data Wrangling**

```{r}

pacman::p_load(readr, dplyr, tidyr, stringr, lubridate,terra,spatstat,
               sf, tmap, units, httr, jsonlite, tidyverse,rvest )

```

### **2.1 Importing ACRA Data (Aspatial)**

```{r}
folder_path <- "data/aspatial/acra"
file_list <- list.files(path=folder_path,
                        pattern = "^ACRA*.*\\.csv$",
                        full.names = TRUE)

acra_data <- file_list %>%
  map_dfr(read_csv)
```

**Saving ACRA data**

```{r}
write_rds(acra_data,
          "data/rds/acra_data.rds") 
```

**Tidying ACRA data (56123 is for bubble tea shops)**

```{r}
### select target businesess, derive year, tidy postal code for geo coding and...
biz_56123 <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code == 56123) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date, label = TRUE, abbr = TRUE)) %>%
  mutate(postal_code = str_pad(postal_code, width = 6, side = "left", pad = "0")) %>%
   filter((YEAR == 2024 & MONTH_NUM >= 1) | (YEAR == 2025 & MONTH_NUM <= 6))

```

```{r}

#|eval:false
#|echo:false
write_rds(biz_56123,
"data/rds/biz_56123.rds")

```

```{r}

biz_56123 <-
read_rds("data/rds/biz_56123.rds")

```

**Geocoding (for finding X,Y coordinates)**

```{r}
postcodes <- unique(biz_56123$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"  ##API without paying

found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )


###converting to data frame
res <- GET(url, query = query)
json <- content(res)

if (json$found != 0) {
  df <- as.data.frame(json$results, stringsAsFactors = FALSE)
  df$input_postcode <- pc
  found <- bind_rows(found, df)
  } else {
  not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
}

```

```{r}
found <- found %>%
  select(1:10)
```

Appending the location information (join it back with 56123 right one with new geo_coded postal_code

2 columns in meter, 2 columns in Lon and Lat

```{r}
biz_56123_join = biz_56123 %>%
   left_join(found,
             by = c('postal_code' = 'POSTAL'))
```

```{r}
write_rds(biz_56123_join, "data/rds/biz_56123_join.rds")
```

**Converting into SF data frame**

```{r}

biz_56123_sf <- st_as_sf(biz_56123_join, 
                         coords = c("X","Y"),
                         crs=3414) 
class(biz_56123_sf)
plot(st_geometry(biz_56123_sf))
```

```{r}
# case-insensitive regex for trailing street-type words (and anything after them)
suffix_re <- "(?i)\\s+(ROAD|RD|STREET|ST|AVENUE|AVE|DRIVE|DR|LANE|LN|CLOSE|CL|CRESCENT|CRES|TERRACE|TERR|WALK|PLACE|PL|PARK|PK|HEIGHTS|HTS|LINK|VIEW|WAY|COURT|CT|CIRCUIT|CIR|RISE|GROVE|GARDENS|GDN)(\\s+.*)?$"

# create Zone from street_name (fallback to STREET_NAME if needed)
biz_56123_sf <- biz_56123_sf %>%
  mutate(
    street_raw = coalesce(street_name),
    Zone = street_raw %>%
      str_replace_all("_", " ") %>%  # tidy underscores
      str_squish() %>%               # trim extra spaces
      str_remove(suffix_re) %>%      # drop trailing street-type
      str_squish() %>%
      str_to_upper()                 # <-- UPPER CASE
  )

# quick check
biz_56123_sf %>% select(street_raw, Zone) %>% head()
```

```{r}
write_rds(biz_56123_sf, "data/rds/biz_56123_sf.rds")
```

### **2.2 Importing Geospatial Data**

```{r}
mpsz_sf <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>%
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
mpsz_sf <- mpsz_sf %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
###testing hands on 2.2
mpsz_cl <- mpsz_sf %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS",
         PLN_AREA_N != "NORTH-EASTERN ISLANDS")
```

```{r}
write_rds(mpsz_cl, 
          "data/rds/mpsz_cl.rds")
```

```{r}
st_crs(mpsz_cl) == st_crs(biz_56123_sf)
```

#### **Mapping the geospatial data sets**

```{r}
ggplot() +
  geom_sf(data = mpsz_cl, fill = "gray", color = "black", size = 0.2) +
  geom_sf(data = biz_56123_sf, color = "black", size = 1.2, alpha = 0.7) +
  labs(title = "Food Kiosks Across Singapore Subzones",
       subtitle = "All layers aligned to EPSG:3414 (SVY21)",
       caption = "Source: URA Master Plan & Data.gov.sg") +
  theme_minimal()
```

### **2.3 Importing MRT Data**

```{r}
mrt <- st_read("data/geospatial/LTAMRTStationExitKML.kml") %>%
  st_transform(crs = 3414)
```

```{r}
#|eval:false
#|echo:false
write_rds(mrt,
"data/rds/mrt.rds")

```

```{r}
mrt<-
read_rds("data/rds/mrt.rds")
```

#### **Mapping the geospatial data sets**

```{r}
ggplot() +
  geom_sf(data = mpsz_cl, fill = "gray", color = "black", size = 0.2) +
  geom_sf(data = mrt, color = "green", size = 1.2, alpha = 0.7) +
  labs(title = "MRT Exit Across Singapore Subzones",
       subtitle = "All layers aligned to EPSG:3414 (SVY21)",
       caption = "Source: URA Master Plan & Data.gov.sg") +
  theme_minimal()
```

### **2.4 Joining ARCA data and Geospatial data**

```{r}
#####will do later
```

## **3 Objective: Clustering Analysis**

```{r}
biz_56123_ppp <- as.ppp(biz_56123_sf)

```

```{r}
class(biz_56123_ppp)
```

```{r}
summary(biz_56123_ppp)
```

### **Creating *owin* object and combining point events**

```{r}
sg_owin <- as.owin(mpsz_cl)
class(sg_owin)
```

```{r}
biz56123_sg_ppp = biz_56123_ppp [sg_owin]
```

#### **3.1aClark-Evan Test without CSR for Nearest Neighbour Analysis**

```{r}
clarkevans.test(biz56123_sg_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```

#### **Statistical conclusion**

-   The test statistic **R = 0.514** (well **below 1**) and the **p-value \< 2.2e-16**.

    The food kioks locations are **much closer together than we’d expect by random chance**—i.e., they are **strongly clustered**, not evenly spread out.

#### **3.1bClark-Evan Test with CSR for Nearest Neighbour Analysis**

```{r}
clarkevans.test(biz56123_sg_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                method="MonteCarlo",
                nsim=99)

```

#### **Statistical conclusion**

-   Clark–Evans NN test: **R = 0.514** (well below 1), **p \< 2.2×10⁻¹⁶** → reject CSR; locations are **significantly clustered**.

    Food kiosks are **much closer together** than random—i.e., **strongly clustered**, not evenly spread across Singapore.

#### **3.2 Kernel Density Method**

```{r}
biz56123sg_ppp_km <- rescale.ppp(biz56123_sg_ppp, 1000, "km")
```

```{r}
kde_biz56123_km <- density(biz56123sg_ppp_km,
                              sigma=bw.diggle,
                              edge=TRUE,
                              kernel="gaussian")
```

```{r}
plot(kde_biz56123_km)
```

```{r}
summary(kde_biz56123_km)
```

#### Statistical Takeaways

-   The **integral ≈ 482.3** matches the number of kiosks → diagnostic check passed (KDE surface conserves counts).

<!-- -->

-    The **mean intensity ≈ 7.2×10⁻⁷ per m²** ≈ **0.72 kiosks per km²** on average over the study window—so most of Singapore is low-intensity with a few concentrated pockets.

<!-- -->

-   **Mean intensity:** **0.720 kiosks/km²** → the island-wide baseline is low; activity is concentrated in a few places.

<!-- -->

-    **Peak intensity:** **\~91.5 kiosks/km²** → very small areas achieve extremely high density, confirming **tight hotspots**.

-   Hotspots align with **MRT interchanges, mall corridors, CBD/town centres**; peripheral/industrial/green areas are cold.

**Business Communications**

-   **Retail gravity rules**: kiosks concentrate where **footfall and delivery access** are highest (CBD/malls/interchanges).

<!-- -->

-    **Saturation risk**: the same nodes show **over-concentration**—new kiosks here will face **tight competition and cannibalisation**.

<!-- -->

-    **Opportunity**: identify **underserved subzones** (low KDE but reasonable demand/transport) as candidate areas for **decentralised growth**.


## **4 Objective: Examine Infrastructure Proximity**
