---
title: "Take-Home Exercise 1: Hotspots on the Go: Clustering of Food Kiosks Across Singapore, 2024–Jun 2025"
author: "Pwint Phoo Thaw"
date: "September 12, 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **Problem Statement**

From **Jan 2024 to Jun 2025**, Singapore saw rapid growth of **food kiosks (SSIC 56123)**—small, takeaway-focused outlets with little or no seating. Their expansion around MRT nodes and malls can reshape micro-retail patterns, affect pedestrian space and curb logistics, and create pockets of under- or over-supply. This study uses ACRA registrations to reveal **where** and **when** kiosks emerge, how strongly they **cluster**, how activity **evolves over time**, and how locations relate to key **urban infrastructure**, to inform **evidence-based retail and land-use planning** in Singapore.

## **1. Objectives**

-   **Assess Clustering \***\
    Use spatial statistics (KDE, Ripley’s K/L or O-ring) to test whether kiosks are clustered or dispersed and at what spatial scales.
-   **Map Spatio-Temporal Distribution\***\
    Track when and where new 56123 kiosks entered (2024–Jun 2025) and highlight hotspot areas by month and subzone.
-   **Examine Infrastructure Proximity \***\
    Test whether kiosks concentrate near **MRT stations, malls, and schools** (buffer analysis; simple statistical tests).
-   **Identify Temporal Waves**\
    Apply spatio-temporal KDE (STKDE) to reveal waves of kiosk entry (e.g., CBD/malls first, then diffusion to heartlands).
-   **Derive Planning Insights**\
    Translate findings into implications for **zoning/tenant-mix**, **transit-oriented clustering management**, curb/logistics planning, and **SME support**.

## **2. Data Wrangling**

```{r}

pacman::p_load(readr, dplyr, tidyr, stringr, lubridate,terra,spatstat,
               sf, tmap, units, httr, jsonlite, tidyverse,rvest, ggthemes, ployly, ggplot2)

```

### **2.1 Importing ACRA Data (Aspatial)**

```{r}
folder_path <- "data/aspatial/acra"
file_list <- list.files(path=folder_path,
                        pattern = "^ACRA*.*\\.csv$",
                        full.names = TRUE)

acra_data <- file_list %>%
  map_dfr(read_csv)
```

**Saving ACRA data**

```{r}
write_rds(acra_data,
          "data/rds/acra_data.rds") 
```

**Tidying ACRA data (56123 is for bubble tea shops)**

```{r}
### select target businesess, derive year, tidy postal code for geo coding and...
biz_56123 <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code == 56123) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date, label = TRUE, abbr = TRUE)) %>%
  mutate(postal_code = str_pad(postal_code, width = 6, side = "left", pad = "0")) %>%
   filter((YEAR == 2024 & MONTH_NUM >= 1) | (YEAR == 2025 & MONTH_NUM <= 6))

```

```{r}

#|eval:false
#|include: false
#|echo:false
write_rds(biz_56123,
"data/rds/biz_56123.rds")

```

```{r}

biz_56123 <-
read_rds("data/rds/biz_56123.rds")

```

**Geocoding (for finding X,Y coordinates)**

```{r}
postcodes <- unique(biz_56123$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"  ##API without paying

found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )


###converting to data frame
res <- GET(url, query = query)
json <- content(res)

if (json$found != 0) {
  df <- as.data.frame(json$results, stringsAsFactors = FALSE)
  df$input_postcode <- pc
  found <- bind_rows(found, df)
  } else {
  not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
}

```

```{r}
found <- found %>%
  select(1:10)
```

Appending the location information (join it back with 56123 right one with new geo_coded postal_code

2 columns in meter, 2 columns in Lon and Lat

```{r}
biz_56123_join = biz_56123 %>%
   left_join(found,
             by = c('postal_code' = 'POSTAL'))
```

```{r}
write_rds(biz_56123_join, "data/rds/biz_56123_join.rds")
```

**Converting into SF data frame**

```{r}

biz_56123_sf <- st_as_sf(biz_56123_join, 
                         coords = c("X","Y"),
                         crs=3414) 
class(biz_56123_sf)

```

```{r}

#|eval:false
#|include: false
#|echo:false
# case-insensitive regex for trailing street-type words (and anything after them)
suffix_re <- "(?i)\\s+(ROAD|RD|STREET|ST|AVENUE|AVE|DRIVE|DR|LANE|LN|CLOSE|CL|CRESCENT|CRES|TERRACE|TERR|WALK|PLACE|PL|PARK|PK|HEIGHTS|HTS|LINK|VIEW|WAY|COURT|CT|CIRCUIT|CIR|RISE|GROVE|GARDENS|GDN)(\\s+.*)?$"

# create Zone from street_name (fallback to STREET_NAME if needed)
biz_56123_sf <- biz_56123_sf %>%
  mutate(
    street_raw = coalesce(street_name),
    Zone = street_raw %>%
      str_replace_all("_", " ") %>%  # tidy underscores
      str_squish() %>%               # trim extra spaces
      str_remove(suffix_re) %>%      # drop trailing street-type
      str_squish() %>%
      str_to_upper()                 # <-- UPPER CASE
  )

# quick check
biz_56123_sf %>% select(street_raw, Zone) %>% head()
```

```{r}
write_rds(biz_56123_sf, "data/rds/biz_56123_sf.rds")
```

### **2.2 Importing Geospatial Data**

```{r}
mpsz_sf <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>%
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
mpsz_sf <- mpsz_sf %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}

mpsz_cl <- mpsz_sf %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS",
         PLN_AREA_N != "NORTH-EASTERN ISLANDS")
```

```{r}
write_rds(mpsz_cl, 
          "data/rds/mpsz_cl.rds")
```

```{r}
st_crs(mpsz_cl) == st_crs(biz_56123_sf)
```

#### **Mapping the geospatial data sets**

```{r}
ggplot() +
  geom_sf(data = mpsz_cl, fill = "gray", color = "black", size = 0.2) +
  geom_sf(data = biz_56123_sf, color = "black", size = 1.2, alpha = 0.7) +
  labs(title = "Food Kiosks Across Singapore Subzones",
       subtitle = "All layers aligned to EPSG:3414 (SVY21)",
       caption = "Source: URA Master Plan & Data.gov.sg") +
  theme_minimal()
```

### **2.3 Importing MRT Data**

```{r}
mrt <- st_read("data/geospatial/LTAMRTStationExitKML.kml") %>%
  st_transform(crs = 3414)
```

```{r}
#|eval:false
#|echo:false
write_rds(mrt,
"data/rds/mrt.rds")

```

```{r}
mrt<-
read_rds("data/rds/mrt.rds")
```

#### **Mapping the geospatial data sets**

```{r}
ggplot() +
  geom_sf(data = mpsz_cl, fill = "gray", color = "black", size = 0.2) +
  geom_sf(data = mrt, color = "green", size = 1.2, alpha = 0.7) +
  labs(title = "MRT Exit Across Singapore Subzones",
       subtitle = "All layers aligned to EPSG:3414 (SVY21)",
       caption = "Source: URA Master Plan & Data.gov.sg") +
  theme_minimal()
```

## **3 Objective: Clustering Analysis**

```{r}
biz_56123_ppp <- as.ppp(biz_56123_sf)

```

```{r}
class(biz_56123_ppp)
```

```{r}
#| eval: false
summary(biz_56123_ppp)
```

### **Creating *owin* object and combining point events**

```{r}
sg_owin <- as.owin(mpsz_cl)
class(sg_owin)
```

```{r}
biz56123_sg_ppp = biz_56123_ppp [sg_owin]
```

#### **3.1 Clark-Evan Test with CSR for Nearest Neighbour Analysis**

```{r}
clarkevans.test(biz56123_sg_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                method="MonteCarlo",
                nsim=99)

```

#### **Statistical conclusion**

-   Clark–Evans NN test: **R = 0.514** (well below 1), **p \< 2.2×10⁻¹⁶** → reject CSR; locations are **significantly clustered**.

    Food kiosks are **much closer together** than random—i.e., **strongly clustered**, not evenly spread across Singapore.

#### **3.2 Kernel Density Method**

```{r}
biz56123sg_ppp_km <- rescale.ppp(biz56123_sg_ppp, 1000, "km")
```

```{r}
kde_biz56123_km <- density(biz56123sg_ppp_km,
                              sigma=bw.diggle,
                              edge=TRUE,
                              kernel="gaussian")
```

#### **Working with different automatic badwidth methods**

```{r}
bw.CvL(biz56123sg_ppp_km )
```

```{r}
bw.scott(biz56123sg_ppp_km)
```

```{r}
bw.ppl(biz56123sg_ppp_km)
```

```{r}
bw.diggle(biz56123sg_ppp_km)
```

```{r}
kde_biz56123sg.ppl <- density(biz56123sg_ppp_km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_biz56123_km, main = "bw.diggle")
plot(kde_biz56123sg.ppl, main = "bw.ppl")
```

#### **Plotting cartographic quality KDE map**

```{r}
kde_biz56123sg_bw_terra <- rast(kde_biz56123_km)
class(kde_biz56123sg_bw_terra)
```

```{r}
kde_biz56123sg_bw_terra
```

```{r}
crs(kde_biz56123sg_bw_terra) <- "EPSG:3414"
kde_biz56123sg_bw_terra
```

```{r}
tm_shape(kde_biz56123sg_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) +
  tm_compass() +
  tm_layout(scale = 1.0)
```

**Insights:**

-   Clear **hotspots at major commercial/commuter nodes**—the **Downtown Core/CBD** and several **regional town centres** (e.g., Jurong East/West, Ang Mo Kio–Bishan, Sembawang, Punggol–Sengkang).

-   **Linear belts** of higher intensity align with **MRT corridors/interchange areas**, suggesting transit-oriented siting.

-   Peaks are **highly localized** (small bright cells) while the **mean surface is low**, indicating concentrated clusters rather than broad uniform coverage—**consistent with Clark-Evans result (R≈0.51)** that showed strong clustering.

#### **3.3 First Order SPPA at the Planning Subzone Level (Sembawang, Ang Mo Kio & Bishan, Punggol & Sengkang, Downtown Core, Jurong West & Jurong East)**

```{r}
sbw <- mpsz_cl %>%
  filter(PLN_AREA_N %in% c( "SEMBAWANG"))

amk_bs <- mpsz_cl %>%
  filter(PLN_AREA_N %in% c("ANG MO KIO", "BISHAN"))

pg_sk<- mpsz_cl %>%
  filter(PLN_AREA_N %in% c("PUNGGOL","SENGKANG"))

dc<- mpsz_cl %>%
  filter(PLN_AREA_N %in% c("DOWNTOWN CORE"))

jw_je <- mpsz_cl %>%
  filter(PLN_AREA_N %in% c("JURONG WEST", "JURONG EAST"))
```

```{r}
sbw_owin = as.owin(sbw)
amk_bs_owin = as.owin(amk_bs)
pg_sk_owin = as.owin(pg_sk)
jw_je_owin = as.owin(jw_je)
dc_owin = as.owin(dc)
```

```{r}
biz56123_sbw_ppp = biz_56123_ppp [sbw_owin]
biz56123_amk_bs_ppp = biz_56123_ppp [amk_bs_owin]
biz56123_pg_sk_ppp = biz_56123_ppp [pg_sk_owin]
biz56123_jw_je_ppp = biz_56123_ppp [jw_je_owin]
biz56123_dc_ppp = biz_56123_ppp [dc_owin]
```

```{r}
biz56123_sbw_ppp.km = rescale.ppp(biz56123_sbw_ppp, 1000, "km")
biz56123_amk_bs_ppp.km = rescale.ppp(biz56123_amk_bs_ppp, 1000, "km")
biz56123_pg_sk_ppp.km = rescale.ppp(biz56123_pg_sk_ppp, 1000, "km")
biz56123_jw_je_ppp.km = rescale.ppp(biz56123_jw_je_ppp, 1000, "km")
biz56123_dc_ppp.km = rescale.ppp(biz56123_dc_ppp, 1000, "km")
```

```{r}
par(mfrow=c(2,3))
plot(unmark(biz56123_sbw_ppp.km ), 
  main="Sembawang")
plot(unmark(biz56123_amk_bs_ppp.km), 
  main="Ang Mo Kio & Bishan")
plot(unmark(biz56123_pg_sk_ppp.km), 
  main="Punggol & Sengkang")
plot(unmark(biz56123_jw_je_ppp.km), 
  main="Jurong West & Jurong East")
plot(unmark(biz56123_dc_ppp.km), 
  main="Downtown Core")
```

#### **Clark and Evans Test**

```{r}
clarkevans.test(biz56123_sbw_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

#### **Computing KDE surfaces by planning area**

```{r}


ppps <- list(
  "Sembawang"                 = biz56123_sbw_ppp.km,
  "Ang Mo Kio & Bishan"       = biz56123_amk_bs_ppp.km,
  "Punggol & Sengkang"        = biz56123_pg_sk_ppp.km,
  "Jurong West & Jurong East" = biz56123_jw_je_ppp.km,
  "Downtown Core"             = biz56123_dc_ppp.km
     )

sigma_km <- 0.30
ims  <- lapply(ppps, function(pp) density(pp, sigma = sigma_km, edge = TRUE, kernel = "gaussian"))
zmax <- max(sapply(ims, function(im) max(im$v, na.rm = TRUE)))

par(mfrow = c(2,3), mar = c(1.2, 1.2, 3.2, 5.2))

# ensure zmax exists (common color scale)
if (!exists("zmax")) {
  zmax <- max(sapply(ims, function(im) max(im$v, na.rm = TRUE)))
}

for (i in seq_along(ims)) {
  plot(ims[[i]],
       main    = "",                 # we'll add title() after
       xlab    = "", ylab = "",
       axes    = FALSE, frame.plot = FALSE,
       ribbon  = TRUE,
       zlim    = c(0, zmax),
       ribargs = list(main = expression("Kiosks per km"^2)))  # legend label

  # 2-line panel title: top = unit, bottom = area name
  ttl <- bquote(atop(" ", .(names(ppps)[i])))
  title(ttl, cex.main = 1.05, line = 1.2)
}
```

```{r}
#| eval: false
#| include: false
#| echo: false

#original

ppps <- list(
  "Sembawang"                 = biz56123_sbw_ppp.km,
  "Ang Mo Kio & Bishan"       = biz56123_amk_bs_ppp.km,
  "Punggol & Sengkang"        = biz56123_pg_sk_ppp.km,
  "Jurong West & Jurong East" = biz56123_jw_je_ppp.km,
  "Downtown Core"             = biz56123_dc_ppp.km
     )

sigma_km <- 0.30
ims  <- lapply(ppps, function(pp) density(pp, sigma = sigma_km, edge = TRUE, kernel = "gaussian"))
zmax <- max(sapply(ims, function(im) max(im$v, na.rm = TRUE)))

#sigma_km <- 0.30
#ims <- lapply(ppps, function(pp) spatstat.core::density(pp, sigma = sigma_km, edge = TRUE, kernel = "gaussian"))
#zmax <- max(sapply(ims, function(im) max(im$v, na.rm = TRUE)))


par(mfrow = c(2,3), mar = c(1.2, 1.2, 2.8, 5.2))
par(mar = c(4,4,3.8,6))
plot(ims[[i]], main = "")
title(ttl, line = 1.2, cex.main = 1.05)

for (i in seq_along(ims)) {
  
  plot(ims[[i]],
       main    = "",
       xlab    = "", ylab = "",
       axes    = FALSE, frame.plot = FALSE,
       ribbon  = TRUE,
       zlim    = c(0, zmax),
       ribargs = list(main = " ")) 

ttl <- bquote(atop("Kiosks per km"^2, .(names(ppps)[i])))
3title(ttl, cex.main = 1.05, line = 1.2)

   }


```

#### **3.4 Second-order Spatial Point Patterns Analysis**

***G Function Estimation (for Sembawang)***

```{r}
set.seed(1234)
```

```{r}

G_SBW = Gest(biz56123_sbw_ppp, correction = "border")
plot(G_SBW, xlim=c(0,500))


```

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of food kiosks at Yishun are randomly distributed.

H1= The distribution of food kiosks at Yishun are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with G-function

```{r}
G_SBW.csr <- envelope(biz56123_sbw_ppp, Gest, nsim = 999)
```

```{r}
plot(G_SBW.csr)
```

***Pairwise L Function(Sembawang, Ang Mo Kio & Bishan, Punggol & Sengkang, Downtown Core, Jurong West & Jurong East)***

```{r}


L_sbw= Lest(biz56123_sbw_ppp, correction = "Ripley")
L_amk_bs= Lest(biz56123_amk_bs_ppp, correction = "Ripley")
L_pg_sk= Lest(biz56123_pg_sk_ppp, correction = "Ripley")
L_jw_je= Lest(biz56123_jw_je_ppp, correction = "Ripley")
L_dc= Lest(biz56123_dc_ppp, correction = "Ripley")

set.seed(1234)
L_sbw.csr <- envelope(biz56123_sbw_ppp, Lest, 
                     nsim = 99, 
                     rank = 1, 
                      global=TRUE)


set.seed(1234)
L_amk_bs.csr <- envelope(biz56123_amk_bs_ppp, Lest, 
                     nsim = 99, 
                     rank = 1, 
                     global=TRUE)


set.seed(1234)
L_pg_sk.csr <- envelope(biz56123_pg_sk_ppp, Lest, 
                     nsim = 99, 
                     rank = 1, 
                     global=TRUE)


set.seed(1234)
L_jw_je.csr <- envelope(biz56123_jw_je_ppp, Lest, 
                     nsim = 99, 
                     rank = 1, 
                     global=TRUE)


set.seed(1234)
L_dc.csr <- envelope(biz56123_dc_ppp, Lest, 
                     nsim = 99, 
                     rank = 1, 
                     global=TRUE)
```

```{r}

library(ggplot2)
library(ggthemes)
library(plotly)

# ----- build the plot data/frame -----
title <- "Pairwise Distance: L function for Sembawang"

L_sbw_csr_df <- as.data.frame(L_sbw.csr)

colour <- c("#0D657D","#ee770d","#D3D3D3")

csr_plot <- ggplot(L_sbw_csr_df, aes(r, obs - r)) +
  geom_line(colour = "#4d4d4d") +
  geom_line(aes(y = theo - r), colour = "red", linetype = "dashed") +
  geom_ribbon(aes(ymin = lo - r, ymax = hi - r),
              alpha = 0.10, colour = "#91bfdb") +
  xlab("Distance r (m)") +
  ylab("L(r) - r") +
  geom_rug(data = L_sbw_csr_df[L_sbw_csr_df$obs >  L_sbw_csr_df$hi,], sides = "b", colour = colour[1]) +
  geom_rug(data = L_sbw_csr_df[L_sbw_csr_df$obs <  L_sbw_csr_df$lo,], sides = "b", colour = colour[2]) +
  geom_rug(data = L_sbw_csr_df[L_sbw_csr_df$obs >= L_sbw_csr_df$lo &
                               L_sbw_csr_df$obs <= L_sbw_csr_df$hi,], sides = "b", colour = colour[3]) +
  theme_tufte() +
  ggtitle(title)

text1 <- "Significant clustering"
text2 <- "Significant segregation"
text3 <- "Not significant clustering/segregation"

# ----- conditional labelling of the rug traces -----
if (nrow(L_sbw_csr_df[L_sbw_csr_df$obs > L_sbw_csr_df$hi,]) == 0) {
  if (nrow(L_sbw_csr_df[L_sbw_csr_df$obs < L_sbw_csr_df$lo,]) == 0) {

    p <- ggplotly(csr_plot, dynamicTicks = TRUE) |>
      style(text = text3, traces = 4) |>
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else if (nrow(L_sbw_csr_df[L_sbw_csr_df$obs >= L_sbw_csr_df$lo &
                               L_sbw_csr_df$obs <= L_sbw_csr_df$hi,]) == 0) {

    p <- ggplotly(csr_plot, dynamicTicks = TRUE) |>
      style(text = text2, traces = 4) |>
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else {

    p <- ggplotly(csr_plot, dynamicTicks = TRUE) |>
      style(text = text2, traces = 4) |>
      style(text = text3, traces = 5) |>
      layout(xaxis = list(rangeslider = list(visible = TRUE)))
  }

} else if (nrow(L_sbw_csr_df[L_sbw_csr_df$obs < L_sbw_csr_df$lo,]) == 0) {

  if (nrow(L_sbw_csr_df[L_sbw_csr_df$obs >= L_sbw_csr_df$lo &
                        L_sbw_csr_df$obs <= L_sbw_csr_df$hi,]) == 0) {

    p <- ggplotly(csr_plot, dynamicTicks = TRUE) |>
      style(text = text1, traces = 4) |>
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else {

    p <- ggplotly(csr_plot, dynamicTicks = TRUE) |>
      style(text = text1, traces = 4) |>
      style(text = text3, traces = 5) |>
      layout(xaxis = list(rangeslider = list(visible = TRUE)))
  }

} else {

  p <- ggplotly(csr_plot, dynamicTicks = TRUE) |>
    style(text = text1, traces = 4) |>
    style(text = text2, traces = 5) |>
    style(text = text3, traces = 6) |>
    layout(xaxis = list(rangeslider = list(visible = TRUE)))
}

p  # print the plotly object


#Hiding the code for the rest

```

**Insights:**

-   The black curve **L(r)−r is \> 0** across the whole range → kiosks are **not** evenly spaced; they aggregate.

-   **Statistically significant clustering:** The curve sits **above the global CSR envelope** from **r ≈ 0 to \~480 m**

-   **Sembawang:** **≈50%** of the tested distance range (≈0–480 m out of ≈0–950 m) shows **significant clustering** of kiosks.

```{r}
#| echo: false 
library(ggplot2)
library(ggthemes)
library(plotly)

title <- "Pairwise Distance: L function for Ang Mo Kio and Bishan"

L_amk_bs_df <- as.data.frame(L_amk_bs.csr)

colour <- c("#0D657D","#ee770d","#D3D3D3")

csr_plot <- ggplot(L_amk_bs_df, aes(x = r, y = obs - r)) +
  geom_line(colour = "#4d4d4d") +
  geom_line(aes(y = theo - r), colour = "red", linetype = "dashed") +
  geom_ribbon(aes(ymin = lo - r, ymax = hi - r),
              alpha = 0.10, fill = "#91bfdb") +
  xlab("Distance r (m)") +
  ylab("L(r) - r") +
  geom_rug(data = L_amk_bs_df[L_amk_bs_df$obs >  L_amk_bs_df$hi,], sides = "b", colour = colour[1]) +
  geom_rug(data = L_amk_bs_df[L_amk_bs_df$obs <  L_amk_bs_df$lo,], sides = "b", colour = colour[2]) +
  geom_rug(data = L_amk_bs_df[L_amk_bs_df$obs >= L_amk_bs_df$lo & L_amk_bs_df$obs <= L_amk_bs_df$hi,],
           sides = "b", colour = colour[3]) +
  theme_tufte() +
  ggtitle(title)

text1 <- "Significant clustering"
text2 <- "Significant segregation"
text3 <- "Not significant clustering/segregation"

# ---- replace rangeslider() with layout(...rangeslider=...) ----
if (nrow(L_amk_bs_df[L_amk_bs_df$obs > L_amk_bs_df$hi,]) == 0) {

  if (nrow(L_amk_bs_df[L_amk_bs_df$obs < L_amk_bs_df$lo,]) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text3, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else if (nrow(L_amk_bs_df[L_amk_bs_df$obs >= L_amk_bs_df$lo &
                              L_amk_bs_df$obs <= L_amk_bs_df$hi,]) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text2, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))
  }

} else if (nrow(L_amk_bs_df[L_amk_bs_df$obs < L_amk_bs_df$lo,]) == 0) {

  if (nrow(L_amk_bs_df[L_amk_bs_df$obs >= L_amk_bs_df$lo &
                       L_amk_bs_df$obs <= L_amk_bs_df$hi,]) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text1, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))
  }

} else {

  p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    layout(xaxis = list(rangeslider = list(visible = TRUE)))
}

p   # render the plotly object

```

**Insights:**

The curve **L(r) − r \> 0** across the domain → kiosks are **clustered rather than evenly spaced**.

**Significant clustering windows (global envelope):**

-   **Short range:** \~**0–250 m** — strong neighbourhood-scale clustering around town-centre/MRT/mall nodes.

-   **Mid range:** \~**560–740 m** — excess pairs at sub-kilometre separation, consistent with **multiple nearby nodes** within one study window.

-   **Ang Mo Kio & Bishan:** **≈35%** of the tested distance range (about **0–250 m** and **560–740 m** out of \~0–1.2 km) shows **significant clustering** of kiosks.

```{r}
#| echo: false 
library(ggplot2)
library(ggthemes)
library(plotly)

title <- "Pairwise Distance: L function for Punggol and Sengkang"

L_pg_sk_df <- as.data.frame(L_pg_sk.csr)

colour <- c("#0D657D","#ee770d","#D3D3D3")

csr_plot <- ggplot(L_pg_sk_df, aes(x = r, y = obs - r)) +
  geom_line(colour = "#4d4d4d") +
  geom_line(aes(y = theo - r), colour = "red", linetype = "dashed") +
  geom_ribbon(aes(ymin = lo - r, ymax = hi - r),
              alpha = 0.10, fill = "#91bfdb") +
  xlab("Distance r (m)") +
  ylab("L(r) - r") +
  geom_rug(data = subset(L_pg_sk_df, obs >  hi), sides = "b", colour = colour[1]) +
  geom_rug(data = subset(L_pg_sk_df, obs <  lo), sides = "b", colour = colour[2]) +
  geom_rug(data = subset(L_pg_sk_df, obs >= lo & obs <= hi), sides = "b", colour = colour[3]) +
  theme_tufte() +
  ggtitle(title)

text1 <- "Significant clustering"
text2 <- "Significant segregation"
text3 <- "Not significant clustering/segregation"

# ---- replace rangeslider() with layout(...rangeslider=...) ----
if (nrow(subset(L_pg_sk_df, obs > hi)) == 0) {

  if (nrow(subset(L_pg_sk_df, obs < lo)) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text3, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else if (nrow(subset(L_pg_sk_df, obs >= lo & obs <= hi)) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text2, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))
  }

} else if (nrow(subset(L_pg_sk_df, obs < lo)) == 0) {

  if (nrow(subset(L_pg_sk_df, obs >= lo & obs <= hi)) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text1, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))
  }

} else {

  p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    layout(xaxis = list(rangeslider = list(visible = TRUE)))
}

p  # render the interactive plot

```

**Insights:**

-   The curve **L(r) − r \> 0 across the whole domain** → kiosks are **aggregated** (not evenly spaced).

-   The observed curve lies **above the CSR envelope from \~10 m to \~110 m** (nearly the entire tested range).\
    ⇒ **strong, statistically significant clustering at very short distances**—kiosks tend to co-locate within the same mall/station complex or adjacent shop lines.

-   **Punggol & Sengkang:** **≈90%** of the tested distance range (**\~10–110 m of \~0–120 m**) shows **significant clustering**.

```{r}
#| echo: false 
library(ggplot2)
library(ggthemes)
library(plotly)

title <- "Pairwise Distance: L function for Jurong West and Jurong East"

L_jw_je_df <- as.data.frame(L_jw_je.csr)

colour <- c("#0D657D","#ee770d","#D3D3D3")

csr_plot <- ggplot(L_jw_je_df, aes(x = r, y = obs - r)) +
  geom_line(colour = "#4d4d4d") +
  geom_line(aes(y = theo - r), colour = "red", linetype = "dashed") +
  geom_ribbon(aes(ymin = lo - r, ymax = hi - r),
              alpha = 0.10, fill = "#91bfdb") +
  xlab("Distance r (m)") +
  ylab("L(r) - r") +
  geom_rug(data = subset(L_jw_je_df, obs >  hi), sides = "b", colour = colour[1]) +
  geom_rug(data = subset(L_jw_je_df, obs <  lo), sides = "b", colour = colour[2]) +
  geom_rug(data = subset(L_jw_je_df, obs >= lo & obs <= hi), sides = "b", colour = colour[3]) +
  theme_tufte() +
  ggtitle(title)

text1 <- "Significant clustering"
text2 <- "Significant segregation"
text3 <- "Not significant clustering/segregation"

# ---- replace rangeslider() with layout(...rangeslider=...) ----
if (nrow(subset(L_jw_je_df, obs > hi)) == 0) {

  if (nrow(subset(L_jw_je_df, obs < lo)) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text3, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else if (nrow(subset(L_jw_je_df, obs >= lo & obs <= hi)) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text2, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))
  }

} else if (nrow(subset(L_jw_je_df, obs < lo)) == 0) {

  if (nrow(subset(L_jw_je_df, obs >= lo & obs <= hi)) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text1, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))
  }

} else {

  p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    layout(xaxis = list(rangeslider = list(visible = TRUE)))
}

p  # render the interactive plot

```

**Insights:**

-   The curve **L(r) − r is positive for almost the entire distance range**, so kiosks are **aggregated** rather than evenly spaced.

<!-- -->

-    The observed curve lies **above the CSR envelope from roughly \~80 m to \~1.85 km**, with only a very short segment near the origin inside the envelope.\
    ⇒ **Strong, statistically significant clustering** from neighbourhood to corridor scale.

-   **Jurong West & Jurong East:** **\~95%** of the tested distance range shows **significant clustering** of kiosks.

```{r}
#| echo: false 
library(ggplot2)
library(ggthemes)
library(plotly)

title <- "Pairwise Distance: L function for Downtown Core"

L_dc_df <- as.data.frame(L_dc.csr)

colour <- c("#0D657D","#ee770d","#D3D3D3")

csr_plot <- ggplot(L_dc_df, aes(x = r, y = obs - r)) +
  geom_line(colour = "#4d4d4d") +
  geom_line(aes(y = theo - r), colour = "red", linetype = "dashed") +
  geom_ribbon(aes(ymin = lo - r, ymax = hi - r),
              alpha = 0.10, fill = "#91bfdb") +
  xlab("Distance r (m)") +
  ylab("L(r) - r") +
  geom_rug(data = subset(L_dc_df, obs >  hi), sides = "b", colour = colour[1]) +
  geom_rug(data = subset(L_dc_df, obs <  lo), sides = "b", colour = colour[2]) +
  geom_rug(data = subset(L_dc_df, obs >= lo & obs <= hi), sides = "b", colour = colour[3]) +
  theme_tufte() +
  ggtitle(title)

text1 <- "Significant clustering"
text2 <- "Significant segregation"
text3 <- "Not significant clustering/segregation"

# ---- replace rangeslider() with layout(...rangeslider=...) ----
if (nrow(subset(L_dc_df, obs > hi)) == 0) {

  if (nrow(subset(L_dc_df, obs < lo)) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text3, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else if (nrow(subset(L_dc_df, obs >= lo & obs <= hi)) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text2, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))
  }

} else if (nrow(subset(L_dc_df, obs < lo)) == 0) {

  if (nrow(subset(L_dc_df, obs >= lo & obs <= hi)) == 0) {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text1, traces = 4) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))

  } else {
    p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      layout(xaxis = list(rangeslider = list(visible = TRUE)))
  }

} else {

  p <- ggplotly(csr_plot, dynamicTicks = TRUE) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    layout(xaxis = list(rangeslider = list(visible = TRUE)))
}

p  # render the interactive plot

```

**Insights:**

-   The curve **L(r) − r stays positive** from \~0 to \~750 m ⇒ kiosks tend to be **aggregated rather than evenly spaced**.

-   However, the observed curve **remains inside the CSR envelope across the whole range** (does **not** exceed the upper envelope).\
    ⇒ **No statistically significant departure from CSR** at the tested alpha (i.e., **clustering is not significant** under the global envelope test).

## **4. Objective: Map Spatio-Temporal Distribution**
