---
title: "Hands-on Exercise 9: Modelling Geographical Accessibility"
author: "Pwint Phoo Thaw"
date-modified: "last-modified"
format: html
execute:
 echo: true
 eval: true
 warning: false
 freeze: true
---

## **9.1 Introduction**

This hands-on exercise is about how to model geographical accessibility by using R’s geospatial analysis packages.

### **9.1.1 Learning Outcome**

-   to computer accessibility measure by using Hansen’s potential model and Spatial Accessibility Measure (SAM); and

-   to visualise the accessibility measures by using tmap and ggplot2 packages.

## **9.2 Data**

Four data sets will be used in this hands-on exercise, they are:

-   `MP14_SUBZONE_NO_SEA_PL`: URA Master Plan 2014 subzone boundary GIS data. This data set is downloaded from data.gov.sg.

-   `hexagons`: A 250m radius hexagons GIS data. This data set was created by using [*st_make_grid()*](https://r-spatial.github.io/sf/reference/st_make_grid.html) of sf package. It is in ESRI shapefile format.

-   `ELDERCARE`: GIS data showing location of eldercare service. [This data](https://data.gov.sg/dataset/eldercare-services) is downloaded from data.gov.sg. There are two versions. One in ESRI shapefile format. The other one in Google kml file format. For the purpose of this hands-on exercise, ESRI shapefile format is provided.

-   `OD_Matrix`: a distance matrix in csv format. There are six fields in the data file.

## **9.3 Getting Started**

```{r}
pacman::p_load(tmap, SpatialAcc, sf, 
               ggstatsplot, reshape2,
               tidyverse)
```

## **9.4 Geospatial Data Wrangling**

### **9.4.1 Importing geospatial data**

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")

```

```{r}
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 
```

```{r}
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 
```

### **9.4.2 Updating CRS information**

```{r}
mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)
```

```{r}
st_crs(mpsz)
```

### **9.4.3 Cleaning and updating attribute fields of the geospatial data**

```{r}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)
```

```{r}
hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

## **9.5 Apsaital Data Handling and Wrangling**

### **9.5.1 Importing Distance Matrix**

```{r}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

### **9.5.2 Tidying distance matrix**

The code chunk below uses *spread()* of **tidyr** package is used to transform the O-D matrix from a thin format into a fat format.

```{r}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```

Currently, the distance is measured in metre because SVY21 projected coordinate system is used. The code chunk below will be used to convert the unit f measurement from metre to kilometre.

```{r}
distmat_km <- as.matrix(distmat/1000)
```

## **9.6 Modelling and Visualising Accessibility using Hansen Method**

### **9.6.1 Computing Hansen’s accessibility**

The code chunk below calculates Hansen’s accessibility using *ac()* of **SpatialAcc** and *data.frame()* is used to save the output in a data frame called `acc_Handsen`.

```{r}

library(SpatialAcc)


D <- distmat_km
if (inherits(D, "units")) D <- units::drop_units(D)
D <- as.matrix(D)

stopifnot(length(hexagons$demand)   == nrow(D))  # origins
stopifnot(length(eldercare$capacity) == ncol(D)) # destinations

acc_Hansen <- SpatialAcc::ac(
  p      = hexagons$demand,      # demand (unused by Hansen but required)
  n      = eldercare$capacity,   # supply
  D      = D,                    # distance matrix
  d0     = max(D, na.rm = TRUE), # dummy; Hansen ignores d0
  power  = 2,
  family = "Hansen"
)


```

```{r}
acc_Hansen <- as.matrix(acc_Hansen)
colnames(acc_Hansen) <- "accHansen"

```

```{r}
acc_Hansen <- as_tibble(acc_Hansen)
```

```{r}
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

one step only

**In-Class Exercise 9**

Trying with power =0.5 to 2.5

::: panel-tabset
## power = 0.5

```{r}

D <- distmat_km
if (inherits(D, "units")) D <- units::drop_units(D)
D <- as.matrix(D)

stopifnot(length(hexagons$demand)   == nrow(D))  # origins
stopifnot(length(eldercare$capacity) == ncol(D)) # destinations

acc_Hansen0.5 <- SpatialAcc::ac(
  p      = hexagons$demand,      # demand (unused by Hansen but required)
  n      = eldercare$capacity,   # supply
  D      = D,                    # distance matrix
  d0     = max(D, na.rm = TRUE), # dummy; Hansen ignores d0
  power  = 0.5,
  family = "Hansen"
)

```

```{r}
acc_Hansen0.5 <- as.matrix(acc_Hansen0.5)
colnames(acc_Hansen0.5) <- "accHansen0.5"

```

```{r}
acc_Hansen0.5 <- as_tibble(acc_Hansen0.5)
```

```{r}
hexagon_Hansen0.5 <- bind_cols(hexagons, acc_Hansen0.5)
```

## power =1

```{r}

D <- distmat_km
if (inherits(D, "units")) D <- units::drop_units(D)
D <- as.matrix(D)

stopifnot(length(hexagons$demand)   == nrow(D))  # origins
stopifnot(length(eldercare$capacity) == ncol(D)) # destinations

acc_Hansen1 <- SpatialAcc::ac(
  p      = hexagons$demand,      # demand (unused by Hansen but required)
  n      = eldercare$capacity,   # supply
  D      = D,                    # distance matrix
  d0     = max(D, na.rm = TRUE), # dummy; Hansen ignores d0
  power  = 1,
  family = "Hansen"
)

```

```{r}
acc_Hansen1 <- as.matrix(acc_Hansen1)
colnames(acc_Hansen1) <- "accHansen1"

```

```{r}
acc_Hansen1 <- as_tibble(acc_Hansen1)
```

```{r}
hexagon_Hansen1 <- bind_cols(hexagons, acc_Hansen1)
```

## power =1.5

```{r}

D <- distmat_km
if (inherits(D, "units")) D <- units::drop_units(D)
D <- as.matrix(D)

stopifnot(length(hexagons$demand)   == nrow(D))  # origins
stopifnot(length(eldercare$capacity) == ncol(D)) # destinations

acc_Hansen1.5 <- SpatialAcc::ac(
  p      = hexagons$demand,      # demand (unused by Hansen but required)
  n      = eldercare$capacity,   # supply
  D      = D,                    # distance matrix
  d0     = max(D, na.rm = TRUE), # dummy; Hansen ignores d0
  power  = 1.5,
  family = "Hansen"
)

```

```{r}
acc_Hansen1.5 <- as.matrix(acc_Hansen1.5)
colnames(acc_Hansen1.5) <- "accHansen1.5"

```

```{r}
acc_Hansen1.5 <- as_tibble(acc_Hansen1.5)
```

```{r}
hexagon_Hansen1.5 <- bind_cols(hexagons, acc_Hansen1.5)
```

## power = 2

```{r}

D <- distmat_km
if (inherits(D, "units")) D <- units::drop_units(D)
D <- as.matrix(D)

stopifnot(length(hexagons$demand)   == nrow(D))  # origins
stopifnot(length(eldercare$capacity) == ncol(D)) # destinations

acc_Hansen2 <- SpatialAcc::ac(
  p      = hexagons$demand,      # demand (unused by Hansen but required)
  n      = eldercare$capacity,   # supply
  D      = D,                    # distance matrix
  d0     = max(D, na.rm = TRUE), # dummy; Hansen ignores d0
  power  = 2,
  family = "Hansen"
)

```

```{r}
acc_Hansen2 <- as.matrix(acc_Hansen2)
colnames(acc_Hansen2) <- "accHansen2"

```

```{r}
acc_Hansen2 <- as_tibble(acc_Hansen2)
```

```{r}
hexagon_Hansen2 <- bind_cols(hexagons, acc_Hansen2)
```

## power = 2.5

```{r}

D <- distmat_km
if (inherits(D, "units")) D <- units::drop_units(D)
D <- as.matrix(D)

stopifnot(length(hexagons$demand)   == nrow(D))  # origins
stopifnot(length(eldercare$capacity) == ncol(D)) # destinations

acc_Hansen2.5 <- SpatialAcc::ac(
  p      = hexagons$demand,      # demand (unused by Hansen but required)
  n      = eldercare$capacity,   # supply
  D      = D,                    # distance matrix
  d0     = max(D, na.rm = TRUE), # dummy; Hansen ignores d0
  power  = 2.5,
  family = "Hansen"
)

```

```{r}
acc_Hansen2.5 <- as.matrix(acc_Hansen2.5)
colnames(acc_Hansen2.5) <- "accHansen2.5"

```

```{r}
acc_Hansen2.5 <- as_tibble(acc_Hansen2.5)
```

```{r}
hexagon_Hansen2.5 <- bind_cols(hexagons, acc_Hansen2.5)
```
:::

### **9.6.2 Visualising Hansen’s accessibility**

**Extracting map extend**

Firstly, we will extract the extend of `hexagons` simple feature data frame by by using *st_bbox()* of **sf** package.

```{r}
mapex <- st_bbox(hexagons)
```

The code chunk below uses a collection of mapping fucntions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.

```{r}
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_polygons(fill = "accHansen", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 10,
                values = "brewer.blues")) +
tm_shape(eldercare) +
  tm_dots(size = 0.3,
          fill = "red",
          col = "black") +
  tm_title("Accessibility to eldercare: Hansen method") +
  tm_layout(frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha = 0.2)
```
