---
title: "In_Class_Ex10:Market Area Analysis for Retail and Service Locations using Huff model"
author: "Pwint Phoo Thaw"
date-modified: "last-modified"
format: html
execute:
 echo: true
 eval: true
 warning: false
 freeze: true
---

## **18.1 Overview**

Huff model is a geospatial analysis method commonly used in market area analysis to predict the probability that a consumer from a specific area will patronize a particular store, based on the store**’s attractiveness and distance relative to competitors.** It calculates a store’s market share by considering factors like size or sales for attractiveness and distance or travel time, with the probability decreasing as distance or competition increases. This tool helps businesses assess potential revenue, define trade areas, and make informed decisions about site selection

### **18.1.1 Learning Outcome**

### **18.1.2 Setting the scene**

A fast-food franchise is a business model in which an individual (the franchisee) acquires the rights to operate a restaurant under an established brand owned by a franchisor. This arrangement typically involves franchise fees and adherence to standardized operational procedures. Key considerations prior to investment include the scale of initial capital outlay, the strength of the brand, and the strategic importance of store location. Capital investment requirements vary significantly, generally ranging from approximately \$200,000 to more than \$2 million. Major global fast-food franchise brands include McDonald’s, Burger King, Pizza Hut, KFC, and Taco Bell.

### **18.1.3 The data**

Four data sets will be used in this hands-on exercise, they are:

-   Geospatial data

    -   STORES_TN, store locations data in ESRI shapefile format.

    -   VILLAGE_TN, village boundary data in ESRI shapefile format.

-   Aspatial data

    -   pop2019.csv, population data of Taichung City in 2019. H_CNT, P_CNT, M_CNT, and F_CNT indicate Number of household, Total population, Male population and Female population respectively.

    -   Store_sales_data.csv provides store sales data, and

## **18.2 Getting Started**

```{r}
pacman::p_load(sf, tidyverse, MCI, tmap, ggstatsplot)

```

## **18.3 Importing Data**

### **18.3.1 Importing geospatial data**

```{r}
stores <- st_read(
  dsn = "data/geospatial",
  layer = "STORES_TC")

```

```{r}
town <- st_read(
  dsn = "data/geospatial",
  layer = "TOWN_TC")
```

```{r}
village <- st_read(
  dsn = "data/geospatial",
  layer = "VILLAGE_TC")
```

```{r}
tmap_mode("view")
tm_shape(town) +
  tm_polygons() +
tm_shape(village) +
  tm_polygons() + 
tm_shape(stores) +
  tm_dots()
```

### **18.3.2 Importing aspatial data**

```{r}
pop2019 <- read_csv(
  "data/aspatial/pop2019.csv")

store_sales <- read_csv(
  "data/aspatial/Store_sales_data.csv")

```

```{r}
summary(store_sales)
```

## **18.4 Data Preparation**

### **18.4.1 Aggregating store sales**

```{r}
store_sales_aggre <- store_sales %>%
  group_by(`Store Code`) %>%
  summarise(
    "Total_Bill" = sum(Bills, na.rm = TRUE),
    "Total_Sales" = sum(Amount, na.rm = TRUE))
```

### **18.4.2 Relational join**

```{r}
stores <- stores %>%
  left_join(store_sales_aggre,
            by = join_by(
              Store_CD == `Store Code`)) %>%
  mutate(Store_size = 100)
```

```{r}
summary(stores)
```

### **18.4.3 Tidying pop2019 data.frame**

```{r}
glimpse(pop2019)
```

```{r}
glimpse(village)
```

Code chunk below will be used to remove the “-” sign from the values.

```{r}
pop2019 <- pop2019 %>%
  mutate(V_ID = str_remove_all(
    V_ID, "-"))
```

```{r}
village <- village %>%
  left_join(pop2019,
            by = join_by(
              VILLCODE == V_ID))
```

### **18.4.4 Check for missing or zero values**

Code chunk below is used to check in any records with missing values.

```{r}
summary(village)
```

```{r}
village <- village %>%
  filter(!is.na(P_CNT)) 
```

```{r}
any(is.na(village$H_CNT))
```

```{r}
any(is.na(village$P_CNT))
```

```{r}
any(is.na(village$M_CNT))
```

```{r}
any(is.na(village$F_CNT))
```

## **18.5 Computing Distance Matrix**

Instead of calculate the distance between the villages and stores, we will calculate the centroid of each village using the code chunk below. Then, compute the distance between the centroids of the villages and the stores.

```{r}
village_center <- st_point_on_surface(village)
distmat <- st_distance(village_center, stores) 
```

```{r}
rownames(distmat) <- village$`VILLCODE`
colnames(distmat) <- stores$`Store_CD`

distmat_long <- distmat %>%
  as.data.frame() %>%
  rownames_to_column(var = "VILLCODE") %>%
  pivot_longer(
    cols = -VILLCODE,
    names_to = "Store_CD",
    values_to = "distance"
  ) %>%
  mutate(distance = as.numeric(distance)) 
```

## **18.6 Preparing Interaction Matrix**

Code chunk below will be used to build the interaction matrix required by MCI package.

```{r}
interaction_matrix <- distmat_long %>%
  left_join(
    stores %>% 
      st_drop_geometry() %>% 
      select(Store_CD, Store_size),
    by = "Store_CD") %>%
  relocate(
    distance, .after = last_col())
```

```{r}
summary(interaction_matrix)
```

## **18.7 Computing Huff model**

```{r}
huff_share <- huff.shares(
  interaction_matrix, "VILLCODE", "Store_CD", 
  "Store_size", "distance")
```

### **18.7.1 Trade area analysis for a single store**

First, let us select market share data of a store by usign *Store_CD* field (i.e. TG).

```{r}
market_share <- huff_share %>%
  filter(Store_CD == "TG")
```

```{r}
Trade_area <- village %>%
  left_join(market_share)

```

```{r}
selected_store <- stores %>%
  filter(Store_CD == "TG")
```

Now, use the code chunk below to plot the trade areas of the selected store.

```{r}
tm_shape(Trade_area) +
  tm_polygons(fill = "p_ij",
              fill.scale = tm_scale_intervals(
                style = "quantile", 
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Patronage Probability")) +
  tm_shape(selected_store) +
    tm_dots(size = 0.3,
            fill = "red",
            col = "black") +
  tm_title("Distribution of Patronage Probability of Store BT by village") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha =0.2) 
```

### **18.7.2 Store sales analysis**

One of the usage of Huff model is to estimate the store sales by using village population.

Firstly, let us append P_CNT field from *village* sf.data.frame to *huff_share* data.frame by using the code chunk below.

```{r}
huff_share_pp <- huff_share %>%
  left_join(village %>% 
      st_drop_geometry() %>% 
      select(VILLCODE, P_CNT),
    by = "VILLCODE") %>%
  relocate(
    distance, .after = last_col())

```

```{r}
huff_total <- shares.total(
  huff_share_pp, "VILLCODE", "Store_CD",
  "p_ij", "P_CNT")
```

```{r}
store_sales_est <- stores %>%
  left_join(huff_total,
            by = join_by(
              "Store_CD" == "suppliers_single"))
```

Now, we can visualise how well the estimate sales as compare to the actual sales by using the code chunk below.

```{r}

store_sales_nongeom <- store_sales_est %>%
  st_drop_geometry()

if (requireNamespace("ggstatsplot", quietly = TRUE)) {
  ggstatsplot::ggscatterstats(
    data = store_sales_nongeom,
    x    = sum_E_j,
    y    = Total_Sales
  )
} else {
  # Fallback: simple scatter + regression line with ggplot2
  library(ggplot2)

  ggplot(store_sales_nongeom,
         aes(x = sum_E_j, y = Total_Sales)) +
    geom_point() +
    geom_smooth(method = "lm", se = TRUE) +
    labs(
      x = "sum_E_j",
      y = "Total Sales",
      title = "Scatterplot with regression line"
    ) +
    theme_minimal()
}

```

```{r}
huff_share_wider <- huff_share %>%
  select(VILLCODE, Store_CD, p_ij) %>%
    pivot_wider(names_from = Store_CD, 
                values_from = p_ij)
```
