---
title: "In-class Exercise 4: Spatial Weight"
author: "Pwint Phoo Thaw"
date: "September 12, 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

```{r}
pacman::p_load(sf, ggstatsplot, tmap, tidyverse, knitr, GWmodel)

```

### Data Preparation

The code chunk below uses [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of **sf** package to import Hunan shapefile into R. The imported shapefile will be **simple features** Object of **sf**.

```{r}
hunan_sf <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

```{r}
#| echo: false
glimpse(hunan_sf)

```

Next, we will import *Hunan_2012.csv* into R by using *read_csv()* of **readr** package. The output is R dataframe class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

```{r}
#| echo: false
glimpse(hunan2012)


```

```{r}
hunan_sf <- left_join(hunan_sf, hunan2012) %>%
  select(1:3, 7, 15, 16, 31, 32)


```

```{r}
hunan_sp <- hunan_sf %>%  
  as_Spatial()
```

#### Geographical analytics solution

#### Cross Validation

```{r}
bw_CV <- bw.gwr(GDPPC ~ 1,
                data = hunan_sp,
                approach = "CV",
                adaptive = FALSE,
                kernel = "bisquare",
                longlat = T)

bw_CV

#better way than looking at Maximum


```

#### AIC

```{r}
bw_AIC <- bw.gwr(GDPPC ~ 1,
                data = hunan_sp,
                approach = "AIC",
                adaptive = FALSE,
                kernel = "bisquare",
                longlat = T)

bw_AIC
#The result 160.5517 in Kilometer
```

#### Summary Statistics

```{r}
#with CV
gwstat_CV <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_CV,
               kernel = "bisquare",
               adaptive = FALSE,
               longlat = T)

```

```{r}
#with AIC
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_AIC,
               kernel = "bisquare",
               adaptive = FALSE,
               longlat = T)

```

#### Extracting SDF data table from gwss object output from gwss()

```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```

####  Combine with Hunan data

```{r}
hunan_gstat <- cbind(hunan_sf,gwstat_df)
#doing column append from gwstat_df to hunan_sf
#use cbind when joining table without unique identifier, only when row sequence are the same

```

```{r}
basemap <- tm_shape(hunan_gstat) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan_gstat, "GDPPC") +
  tm_layout(
    legend.position = c("left", "bottom"))
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

## **Visualising geographically weighted summary statistics**

```{r}
tm_shape(hunan_gstat) +
  tm_fill("GDPPC_LM",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Distribution of geographically weighted mean",
            main.title.position = "center",
            main.title.size = 1.8,
            legend.text.size = 1.2,
            legend.height = 1.50,
            legend.width = 1.50,
            frame = TRUE)
```
