---
title: "Hands-on Exercise 02 Part 2: 2nd Order Spatial Point Patterns Analysis Methods"
author: "Pwint Phoo Thaw"
date-modified: "last-modified"
execute:
  echo: true
  eva: true
  warning: false
  freeze: true
---

## **5.1 Overview**

Second-order spatial point pattern analysis examines the spatial relationships between points in a pattern, specifically focusing on how the presence of one point influences the location of others. It goes beyond simply describing the overall density of points (first-order effects) by investigating clustering, dispersion, or randomness at various spatial scales.

Using appropriate functions of [spatstat](https://cran.r-project.org/web/packages/spatstat/), this hands-on exercise aims to discover the spatial point processes of childecare centres in Singapore.

The specific questions we would like to answer are as follows:

-   are the childcare centres in Singapore randomly distributed throughout the country?

-   if the answer is not, then the next logical question is where are the locations with higher concentration of childcare centres?

## **5.2 The data**

## **5.3 Installing and Loading the R packages**

```{r}
pacman::p_load(sf, terra, spatstat, tmap, tidyverse, rvest)
```


## **5.4 Data Import and Preparation**

The processes of importing and preparing the geospatial data to meet the analysis are similar to Chapter 4. Please refer to the following sub-sections, if necessary:

-   [Importing and Wrangling Geospatial Data Sets](file:///C:/tskam/r4gdsa/_book/chap04.html#importing-and-wrangling-geospatial-data-sets)

-   [Geospatial Data wrangling](file:///C:/tskam/r4gdsa/_book/chap04.html#geospatial-data-wrangling) for preparing geospatial data in ppp object class required by spatstat package.

```{r}
mpsz_sf <- st_read("Data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>% 
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```



```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  if (length(value) == 0) NA_character_ else value
}

```

```{r}
mpsz_sf <- mpsz_sf %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
mpsz_cl <- mpsz_sf %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS",
         PLN_AREA_N != "NORTH-EASTERN ISLANDS")
```

```{r}
write_rds(mpsz_cl, "Data/geospatial/mpsz_cl.rds")
```

```{r}
childcare_sf <- st_read("Data/geospatial/ChildCareServices.kml") %>% 
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 3414)
```

```{r}
st_crs(mpsz_cl) == st_crs(childcare_sf)
```

```{r}
st_crs(mpsz_cl) == st_crs(childcare_sf)
```

```{r}
mpsz_cl <- read_rds("Data/geospatial/mpsz_cl.rds")
childcare_sf <- st_read("Data/geospatial/ChildCareServices.kml") %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = st_crs(mpsz_cl))  # Ensure same CRS
```

```{r}

# Create the map
ggplot() +
  geom_sf(data = mpsz_cl, fill = "gray", color = "black", size = 0.2) +
  geom_sf(data = childcare_sf, color = "black", size = 1.2, alpha = 0.7) +
  labs(title = "Childcare Services Across Singapore Subzones",
       subtitle = "All layers aligned to EPSG:3414 (SVY21)",
       caption = "Source: URA Master Plan & Data.gov.sg") +
  theme_minimal()
```

**Interactive mapping with tmap()**

```{r}
tmap_mode("view")
tm_shape(childcare_sf) +
  tm_dots()
```

Switching back to plot

```{r}
tmap_mode("plot")
```

```{r}
childcare_ppp <- as.ppp(childcare_sf)
```

```{r}
class(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```

```{r}
sg_owin <- as.owin(mpsz_cl)
```

```{r}
class(sg_owin)
```

```{r}
plot(sg_owin)

```

```{r}
childcareSG_ppp = childcare_ppp[sg_owin]

```

```{r}
childcareSG_ppp
```

```{r}
pg <- mpsz_cl %>%
  filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_cl %>%
  filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_cl %>%
  filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_cl %>%
  filter(PLN_AREA_N == "JURONG WEST")

```

```{r}
par(mfrow=c(2,2))
plot(st_geometry(pg), main = "Punggol")
plot(st_geometry(tm), main = "Tampines")
plot(st_geometry(ck), main = "Choa Chu Kang")
plot(st_geometry(jw), main = "Jurong West")
```

```{r}
pg_owin = as.owin(pg)
tm_owin = as.owin(tm)
ck_owin = as.owin(ck)
jw_owin = as.owin(jw)
```

```{r}
childcare_pg_ppp = childcare_ppp[pg_owin]
childcare_tm_ppp = childcare_ppp[tm_owin]
childcare_ck_ppp = childcare_ppp[ck_owin]
childcare_jw_ppp = childcare_ppp[jw_owin]
```

```{r}
childcare_pg_ppp.km = rescale.ppp(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale.ppp(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale.ppp(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale.ppp(childcare_jw_ppp, 1000, "km")
```

```{r}
par(mfrow=c(2,2))
plot(unmark(childcare_pg_ppp.km), 
  main="Punggol")
plot(unmark(childcare_tm_ppp.km), 
  main="Tampines")
plot(unmark(childcare_ck_ppp.km), 
  main="Choa Chu Kang")
plot(unmark(childcare_jw_ppp.km), 
  main="Jurong West")

```

## **5.5 Second-order Spatial Point Patterns Analysis**

## **5.6 Analysing Spatial Point Process Using G-Function**

The G function measures the distribution of the distances from an arbitrary event to its nearest event. In this section, you will learn how to compute G-function estimation by using [*Gest()*](https://rdrr.io/cran/spatstat/man/Gest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using [*envelope()*](https://rdrr.io/cran/spatstat/man/envelope.html) of **spatstat** package.

### **5.6.1 Choa Chu Kang planning area**

#### 5.6.1.1 Computing G-function estimation

```{r}
set.seed(1234)
```

The code chunk below is used to compute G-function using *Gest()* of **spatat** package.

```{r}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```

#### 5.6.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with G-fucntion

```{r}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
```

```{r}
plot(G_CK.csr)
```

### **5.6.2 Tampines planning area**

#### 5.6.2.1 Computing G-function estimation

```{r}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)
```

#### 5.6.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
```

```{r}
plot(G_tm.csr)
```

## **5.7 Analysing Spatial Point Process Using F-Function**

The F function estimates the empty space function F(r) or its hazard rate h(r) from a point pattern in a window of arbitrary shape. In this section, you will learn how to compute F-function estimation by using [*Fest()*](https://rdrr.io/cran/spatstat/man/Fest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using [*envelope()*](https://rdrr.io/cran/spatstat/man/envelope.html) of **spatstat** package.

### **5.7.1 Choa Chu Kang planning area**

#### 5.7.1.1 Computing F-function estimation

The code chunk below is used to compute F-function using *Fest()* of **spatat** package.

```{r}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

### **5.7.2 Performing Complete Spatial Randomness Test**

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with F-fucntion

```{r}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
```

```{r}
plot(F_CK.csr)
```

### **5.7.3 Tampines planning area**

#### 5.7.3.1 Computing F-function estimation

Monte Carlo test with F-fucntion

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

#### 5.7.3.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
```

```{r}
plot(F_tm.csr)
```

## **5.8 Analysing Spatial Point Process Using K-Function**

K-function measures the number of events found up to a given distance of any particular event. In this section, you will learn how to compute K-function estimates by using [*Kest()*](https://rdrr.io/cran/spatstat/man/Kest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using *envelope()* of spatstat package.

### **5.8.1 Choa Chu Kang planning area**

#### 5.8.1.1 Computing K-fucntion estimate

```{r}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

#### 5.8.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

### **5.8.2 Tampines planning area**

#### 5.8.2.1 Computing K-fucntion estimation

```{r}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

#### 5.8.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

## **5.9 Analysing Spatial Point Process Using L-Function**

In this section, you will learn how to compute L-function estimation by using [*Lest()*](https://rdrr.io/cran/spatstat/man/Lest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using *envelope()* of spatstat package.

### **5.9.1 Choa Chu Kang planning area**

#### 5.9.1.1 Computing L Fucntion estimation

```{r}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

#### 5.9.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value if smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

### **5.9.2 Tampines planning area**

#### 5.9.2.1 Computing L-fucntion estimate

```{r}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

#### 5.9.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below will be used to perform the hypothesis testing.

```{r}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```
